---
title: "PMarquardt_CBallenger_BAKannan_CaseStudy1"
author: "Paul Marquardt"
date: "June 17, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(plyr)
library(ggplot2)
```

### MSDS 6306: Case Study 1 Introduction
The following document represents a case study of craft breweries and the beers they produce across the United States. In this study we will make some general observations with the hope of providing information concerning details of the craft brewery market in the United States. This study does not represent an exhaustive list of craft breweries and their beers, but does contain general information about the overall craft beer market.Craft breweries are growing across the nation, as we can see in the following table.

Table 1: Contains a total count of the number of breweries by their respective state
```{r}
# Question 1. How many breweries are present in each state?
# Read the breweries.csv into a dataframe and assign the variable name brew
brew <- read.csv(file = "breweries.csv", header = TRUE, sep = ",", strip.white = TRUE)
# Load the plyr library
library(plyr)
# Assign the count function against the State column to the variable name st
st <- count(brew, "State")
# Display the st variable output
st
```

There are numerous different types of beer styles. Each style of beer is created using differnt types of ingredients which include barley, malt, hops, and yeast. This also includes cider and lambic which is made from fruit juice. 

Table 2: Lists several different styles of beer and corresponding data including hard cider
```{r}
# Question 2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file. 
# Read the beers.csv and assign it to the beer variable
beer <- read.csv(file = "beers.csv", header = TRUE, sep = ",", strip.white = TRUE)
# Merge the beers and brew data frames into the brewbeer variable
brewbeer <- merge(brew, beer, by.x = "Brew_ID", by.y="Brewery_id")
# Display the first 6 lines of the merged dataset
head(brewbeer)
# Display the last 6 lines of the merged dataset
tail(brewbeer)
``` 

### Question 3. Report the number of NA's in each column. 
In order to determine how many NAs are in each column, we will use the sapply() function.  The function will allow us to sum all true values or 1 that are NA's.  Any values that do not have NA's will be returned as False or 0, thus not being included in the sum.
```{r}
sapply(brewbeer, function(y) sum( length( which( is.na( y ) ) ) ) )
``` 
Analyzing the data, we have a large gap in IBU's of 1005 missing values out of 2410 rows.  There is also missing values in ABV, but small enough to continue with our analysis.

### Question 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare. 
Using the aggregate function, we are about to get the median value for ABV and IBU.  Due to the incomplete data for IBU and ABV, we are ignoring any NA values for the purpose of this study.
```{r}
median_abv_ibu <- aggregate( brewbeer[ , 7:8 ], list( brewbeer$State ), median, na.rm=TRUE )
``` 

Removing any possible NAs from the Median value of ABV.
```{r fig.height=10}
ggplot( median_abv_ibu[!is.na(median_abv_ibu$IBU),], aes( Group.1, ABV ) ) +
    geom_bar(stat="identity") +
    coord_flip()
```

Removing any possible NAs from the Median value of IBU.
```{r fig.height=10}
ggplot( median_abv_ibu[!is.na(median_abv_ibu$IBU),] , aes( Group.1, IBU ) ) +
    geom_bar(stat="identity") +
    coord_flip()
```

### Question 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer? 
In order to identify which state's beer has the most alcohol we need to performt he following steps
* Get the max ABV from each State and Merge it to the original data set
* Perform a logical function to compare the beer list's ABV to the Max ABV
* Sort by ABV to show the top ABV Beers.
```{r}
max_abv_ibu <- aggregate( brewbeer[ , 7:8 ], list( brewbeer$State ), max, na.rm=TRUE )
names( max_abv_ibu ) <- c( 'max_State', 'max_abv', 'max_ibu')
# Merge Into Existing Data Set to do Logical Comparsons
brewbeer <- merge( brewbeer, max_abv_ibu, by.x = "State", by.y = "max_State" )
#Find the Beer that matches the max ABV by its State
highest_abv_by_state <- brewbeer[brewbeer$ABV >= brewbeer$max_abv,c( 1, 3, 5, 7, 9)]
#Filter our NAs
highest_abv_by_state <- highest_abv_by_state[which(!is.na(highest_abv_by_state$State)), ]
highest_abv_by_state[order(highest_abv_by_state$ABV, decreasing=TRUE),]
``` 

Same process was done for IBU
```{r}
#Find the Beer that matches the max IBU by its State
highest_ibu_by_state <- brewbeer[brewbeer$IBU >= brewbeer$max_ibu,c( 1, 3, 5, 8, 9)]
#Filter our NAs
highest_ibu_by_state <- highest_ibu_by_state[which(!is.na(highest_ibu_by_state$State)), ]
highest_ibu_by_state[order(highest_ibu_by_state$IBU, decreasing=TRUE),]
``` 

Reviewing the data, we find that beers with the highest IBU is aligned with the same style accros all the states, IPA.  However, higher ABV beers have a wide variety of styles, but most being Imperial IPAs or Stours.

### Quetion 6. Summary statistics for the ABV variable. 
```{r}

``` 
### Question 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. 
```{r}

``` 
You are welcome to use the ggplot2 library for graphs. Please ignore missing values in your analysis. Make your best judgment of a relationship and EXPLAIN your answer. 
 